Automated topic naming to support cross-project analysis of software maintenance activities

### 2011-01-25 msr/msr2011.tex
todos for viz

\begin{abstract}
% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
Researchers have employed a variety of techniques to extract underlying
topics that relate to software development artifacts.  Typically, these
techniques use unsupervised machine-learning algorithms to suggest
candidate word-lists.  However, word-lists are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
from source control systems such as CVS and BitKeeper.  
These topics are given labels from a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on two large-scale RDBMS projects: MySQL and MaxDB.  The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels relevant to these projects, which provides fresh
insight into their evolving software development activities.
\end{abstract}

@@

One challenge for software maintenance is to identify the reasons
behind software development activities. 
Identifying these reasons provides insight into \emph{why} certain
changes were made. 
Topic modeling is a machine learning technique which creates
multinomial distributions over words from a text corpus. 
The technique infers the hidden structure of a corpus using posterior
inference: the probability of the hidden structure given the data. 
Topic models are useful in software maintenance because they summarize
the key concepts in a corpus, such as source code, commit comments, or
mailing-list messages, by identifying statistically co-occurring words. 
Among other uses, it can quickly give developers an overview of where
activity occurred, and gives managers or maintainers a sense of
project history. 
For example, a topic extracted from a set of commit logs might be
\emph{ ``listener change remove add fire''}. 
A major challenge for topic modeling is to accurately identify the
summary label for a topic; in the previous case, perhaps
\emph{``event-handling''} would be appropriate. 
Current approaches are manual: they rely on human expertise to
recognize the appropriate label. 
Furthermore, to date the results of topic modeling are project
specific, and not generalizable. 
This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.

@@

% harp on data mining more
In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
% I was told if you're going to use a paren, you don't need to say it.
There is a series of standards on NFRs, such as ISO9126~\cite{iso9126}, specifically intended to apply to projects of varying types.
These standards indicate our approach is possible. %, if perhaps not universally accepted).
% NFRs are prevalent in almost all software systems.
% In this paper we will show that NFR related topics occur quite frequently.
%XXXXXX this sentence is confusing
%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
In particular, we use software quality models to generalize topics across projects. 
This allows us to mine software systems with the intention of comparing their NFR-related development topics.
% in order to achieve generalizable, cross project data mining results.

@@

Concrete applications of topics and \emph{labelled topic extraction}
range from project dashboards to annotating software artifacts such as revisions and bug reports with
NFR-related tags.
Project dashboard are typically employed by managers and are used to
provide quick summaries of the effort put into a software project. In this case, 
labelled topic extraction would allow managers to track effort
related to topics, in this paper, NFRs such as usability.
These techniques also allow the annotation of commit comments
 and other software artifacts with NFRs. 
This would enable querying of bug reports and artifacts by relevant NFRs. 
For instance one could look
for \emph{usability}-relevant revisions and bug reports.

@@

This paper makes the following contributions: 
\begin{itemize}
\item We introduce the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels. 
\item We evaluate an unsupervised method of automated topic labelling (word-lists);
%\item Our unsupervised method can be applied to projects without training;
\item We evaluate a supervised method of automatically labelling topics with a single NFR (machine learning);
\item We evaluate a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
\item We provide a method of cross-project analysis via topic labelling;
\item We apply these techniques to visualize NFRs over time, and to analyze maintenance activities.
\end{itemize}

@@

The idea of extracting higher-level \emph{concerns}, also known as \emph{concepts}, \emph{aspects} or \emph{requirements}, has been approached from documentation-based and repository-based perspectives.

@@

% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
In Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
The labels are reasonable enough, but still require an expert in the field to determine them. 
Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called `Oracle' in the MySQL case, since Oracle owns MySQL. 
We do this differently.
%This paper does this differently. 
First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.

@@

\section{Understanding software maintenance activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
%In order to test  our ROC values produced acceptable results, we now turn to
%\emph{applying} those techniques to understand software maintenance
%activities. 
We evaluated two research questions using the ground truth data that we
annotated by hand:
\begin{enumerate}
\item Do label frequencies change over time? Is a certain quality of more interest at one point in the life-cycle than some other? 
\item  Do the different projects differ in their relative topic interest? Is a particular quality more important to one project than the other projects?  
\end{enumerate}

@@

%This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
This analysis allows to address the questions raised in Section \ref{sec:analysis}:

@@

\section{Discussion}
\label{sec:limit}

>>>>

\begin{abstract}
% %\todo[inline]{Topics and LDA sooner, Topics/concept location is a known
Researchers have employed a variety of techniques to extract underlying
topics that relate to software development artifacts.  Typically, these
techniques use unsupervised machine-learning algorithms to suggest
candidate word-lists.  However, word-lists are
difficult to 
interpret in the absence of meaningful summary labels.  Current topic modeling 
techniques assume manual labelling and do not use domain-specific knowledge
to improve, contextualize, or describe results for the developers.  
We propose a solution: \emph{automated labelled topic extraction}.
  Topics are extracted 
using  Latent Dirichlet Allocation (LDA) from commit-log comments recovered
from source control systems such as CVS and BitKeeper.  
These topics are given labels from to a generalizable
cross-project taxonomy, consisting of non-functional
requirements. Our approach was evaluated with experiments and case studies
on two large-scale RDBMS projects: MySQL and MaxDB.  The case studies
show that labelled topic extraction can produce appropriate,
context-sensitive labels relevant to these projects, which provides fresh
insight into their evolving software development activities.
\end{abstract}

@@

One challenge for software maintenance is to identify the reasons behind software development activities. Identifying these reasons provides insight into \emph{why} certain changes were made. Topic modeling is a machine learning technique which creates multinomial distributions over words from a text corpus. The technique infers the hidden structure of a corpus using posterior inference (the probability of the hidden structure given the data). Topic models are useful in software maintenance because they summarize the key concepts in a corpus, such as source code, commit comments, or email, by identifying statistically co-occurring words. Among other uses, it can quickly give developers an overview of where activity occurred, and gives managers or maintainers a sense of project history. For example, a topic extracted from a set of commit logs might be \emph{ ``listener change remove add fire''}.  A major challenge for topic modeling is to accurately identify the summary label for a topic (in the previous case, perhaps \emph{``event-handling''}). 
Current approaches are manual: they rely on human expertise to recognize the appropriate label. Furthermore, to date the results of topic modeling are project specific, and not generalizable. This paper presents an approach, \emph{labelled topic extraction}, which addresses these two problems.

@@

% harp on data mining more
In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. 
% I was told if you're going to use a paren, you don't need to say it.
There is a series of standards on NFRs, such as \cite{iso9126}, specifically intended to apply to projects of varying types.
These standards indicate our approach is possible. %, if perhaps not universally accepted).
% NFRs are prevalent in almost all software systems.
% In this paper we will show that NFR related topics occur quite frequently.
%XXXXXX this sentence is confusing
%This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. 
In particular, we use software quality models to generalize topics across projects. 
This allows us to mine software systems with the intention of comparing their NFR-related development topics.
% in order to achieve generalizable, cross project data mining results.

@@

Concrete applications of topics and \emph{labelled topic extraction}
range from project dashboards to annotating software artifacts such as revisions and bug reports with
NFR-related tags.
Project dashboard are typically employed by managers and are used to
provide quick summaries of the effort put into a software project. In this case, 
labelled topic extraction would allow managers to track effort
related to topics, in this paper, NFRs such as usability.
These techniques also allow the annotation of commit comments
 and other software artifacts with NFRs. This would enable querying of bug
reports and artifacts by relevant NFRs. 
For instance one could look
for usability-relevant revisions and bug reports.

@@

This paper makes the following contributions: 
\begin{itemize}
\item We introduce the concept of labelled topic extraction, using a non-functional requirements (NFR) taxonomy for our labels. 
\item We evaluate an unsupervised method of automated topic labelling (word-lists);
%\item Our unsupervised method can be applied to projects without training;
\item We evaluate a supervised method of automatically labelling topics with a single NFR (machine learning);
\item We evaluate a supervised method of labelling topics with multiple NFRs (multi-label machine learning);
\item We provide a method of cross-project analysis via topic labelling;
\item We apply these techniques to visualizing NFRs over time, to analyze maintenance activities.
\end{itemize}

@@

The idea of extracting higher-level `concerns', also known as `concepts', `aspects' or `requirements', has been approached from documentation-based and repository-based perspectives.

@@

% in our previous work~\cite{Hindle09ICSM}, as well as -- CITED a little too often (vanity cites!)
In Baldi et al.~\cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a label accordingly. 
As discussed earlier, given the topic \emph{``listener change remove add fire''}, Baldi et al. would assign the label \emph{event-handling}. 
The labels are reasonable enough, but still require an expert in the field to determine them. 
Furthermore, these labels are project-specific, because they are generated from the data of that project. E.g., we might have a label called `Oracle' in the MySQL case, since Oracle owns MySQL.

This paper does this differently. First of all, we automated the process of naming the topics. Secondly, we labelled topics with project-independent terms, in order to permit cross-project comparison.

@@

\section{Understanding software maintenance activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
%In order to test  our ROC values produced acceptable results, we now turn to
%\emph{applying} those techniques to understand software maintenance
%activities. 
\todo[inline]{more on intro ... why should the reader care about this. It needs to show how the stuff in the intro (applications) is possible}
\todo[inline]{fix section headings}

We evaluated two research questions using the ground truth data that we
annotated by hand:
\begin{enumerate}
\item Do label frequencies change over time? Is a certain quality of more interest at one point in the life-cycle than some other? 
\item  Do the different projects differ in their relative topic interest? Is a particular quality more important to one project than the other projects?  
\end{enumerate}

@@

%This analysis allows to address the questions raised in Section \ref{sec:suplabelling}:
This analysis addresses the questions raised in Section \ref{sec:analysis}:

@@

\todo[inline]{link the actual data, which is interesting, to somethign actionable for a developer or manager}
\section{Discussion}
\label{sec:limit}
